Carbonate chemistry at Ny-Ålesund

### 2021-03-30 awipev-CO2.Rmd
Added K1K2 from AT and CT

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4 = pHinsi(pH = pH_durafet$ph_s_dur, 
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
   )
# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet %>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   #ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_4_mean = mean(ph_s_sf_temp_4, na.rm = TRUE)
                   )
pH_durafet_mean <- pH_durafet %>%
  dplyr::select(date, ph_s_dur_temp_fb, ph_s_dur_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE),
                   #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_4_mean = mean(ph_s_dur_temp_4, na.rm = TRUE)
                   )

@@

# pH calculated (at and ct) at Ferrybox sal and temp
discrete_no_na <- discrete %>%
  dplyr::filter(!is.na(sal_fb)  & !is.na(temp_fb) & !is.na(at) & !is.na(ct))
discrete_is_na <- discrete %>%
  dplyr::filter(is.na(sal_fb)  | is.na(temp_fb) | is.na(at) | is.na(ct))
carb <- carb(15, discrete_no_na$at*1e-6, discrete_no_na$ct*1e-6, S = discrete_no_na$sal_fb, T = discrete_no_na$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
discrete_no_na$phcalc_fb <- carb$pH # pH to compare with durafet pH (at Ferrybox temp and sal)
discrete_no_na$pco2_calc_fb <- carb$pCO2

# create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$phcalc_fb <- NA
discrete_is_na$pco2_calc_fb <- NA
discrete <- bind_rows(discrete_no_na, discrete_is_na) %>% 
  dplyr::arrange(date)
```
```{r read and prepare nutrient data, echo=FALSE, warning=TRUE, message=FALSE}
# File "Ferrybox nutrients 2014-2018.xlsx" provided by 
# Willem van de Poll <w.h.van.de.poll@rug.nl> on 2020-01-16
# nutrients in umol/l
nuts <- readxl::read_excel(paste0(path, "fb_data/Ferrybox nutrients 2014-2018.xlsx"),
                           skip =1) %>% 
  drop_na(date) %>% 
  dplyr::mutate(datetime = as.POSIXct(paste0(date, " 12:00:00")),
                ) %>% # assuming sampling time 12:00
  dplyr::filter(datetime > as.POSIXct("2015-07-09"))
# The nearest times
time_index <- unique(d_all$datetime)
# nuts$closest <- time_index[
#   FNN::knnx.index(time_index, # The values to be queried (big df)
#                   as.vector(nuts$datetime), # The value to find the closest match to (small df)
#                   k = 1)] # The number of closest values to return
nuts <- nuts %>% 
  dplyr::mutate(closest = time_index[
  FNN::knnx.index(time_index, # The values to be queried (big df)
                  as.vector(nuts$datetime), # The value to find the closest match to (small df)
                  k = 1)]) %>%  # The number of closest values to return
  dplyr::select(-date, -datetime) %>% 
  dplyr::rename(NO3NO2 = `NO3+NO2`)
# cleanup and convert in umol/kg
d_all <- left_join(d_all, nuts, by = c("datetime" = "closest")) #%>% 
  #dplyr::select(-closest_datetime)
d_all <- d_all %>% 
  dplyr::mutate(density_fb = seacarb::rho(S = sal_fb, T = temp_fb, P = 0),
                po4 = PO4 / (density_fb/1000),
                nh4 = NH4 / (density_fb/1000),
                no3no2 = NO3NO2 / (density_fb/1000),
                no3 = NO3 / (density_fb/1000),
                si = Si / (density_fb/1000)
  )
```

@@

# pH calculated ferrybox
discrete2merge <- discrete %>%
  #drop_na() %>%
  dplyr::select(closest_datetime, phcalc_fb) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

# durafet pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_dur_temp_fb_mean, ph_s_dur_temp_4_mean) %>%
  dplyr::rename(ph_s_dur_temp_fb = ph_s_dur_temp_fb_mean,
                ph_s_dur_temp_4 = ph_s_dur_temp_4_mean) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

### **Calculated / Durafet pHs**
```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize phcalc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb, temp_fb, phcalc_fb, at)%>%
  dplyr::filter(!is.na(phcalc_fb), !is.na(temp_fb), !is.na(sal_fb))
pH_calc$pH_calc_fb_temp_4 <- pHinsi(pH = pH_calc$phcalc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using mean discrete values
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
mean_discrete <- tibble(mean_ct = mean_ct, mean_at = mean_at, mean_pco2 = mean_pco2, mean_sal = mean_sal, mean_temp = mean_temp)
# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(mean_discrete, digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

@@

### K1K2 using FerryBox pCO2 and at in situ
```{r impact of constants using FerryBox pCO2 and at in situ, echo=FALSE, warning=FALSE, message=FALSE}

@@

  dplyr::select(datetime, ph_s_dur_temp_4, pco2_fb, at, sal_fb, temp_fb)

@@

tmp2 <- tibble(datetime = tmp$datetime, ph_s_dur_temp_4 = tmp$ph_s_dur_temp_4, 

@@

  tidyr::drop_na(ph_s_dur_temp_4)

@@

lueker <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_l))
sulpis <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_s))
papadimitriou <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_p))
millero2002 <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_m))

@@

p1 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 

@@

            xlab = "Spectrophotometric pH measured @ 4°C",

@@

p2 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 

@@

  dplyr::mutate(delta_lueker = tmp2$ph_s_dur_temp_4 - tmp2$pH_l, 
                delta_sulpis = tmp2$ph_s_dur_temp_4 - tmp2$pH_s, 
                delta_papadimitriou = tmp2$ph_s_dur_temp_4 - tmp2$pH_p, 
                delta_millero2002 = tmp2$ph_s_dur_temp_4 - tmp2$pH_m) %>% 

@@

ggsave(filename = "figures/k1k2_all.png", plot = p, width = 25, height = 20, units = "cm")
ggsave(filename = "figures/k1k2_temperature.png", plot = p3, width = 25, height = 20, units = "cm")
```{r save elements for k1k2 report, echo=FALSE, message= FALSE, warning=FALSE}
k1k2 <- list(mean_discrete = mean_discrete, const_choice = const_choice, tmp2 = tmp2,
             errors_pco2_at = errors_pco2_at, diff_table = diff_table, 
             diff_ls_quant = diff_ls_quant)
save(k1k2, file="reports/k1k2_data.RData")

The discrete values of pCO2, salinity and temperature in the FerryBox and in situ AT were used to calculate pH from measured pCO2 and AT using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). Note that this cannot be done with in situ AT and CT because spectrophotometric pH in situ is not at a fixed depth.

The differences induced by the choice of the carbonic acid constants is:

- small for pH (`r round(diff_ls_quant[4, 2], digits = 3)` pH units on average; ; median is `r round(diff_ls_quant[3, 2], digits = 3)`) compared with the uncertainty calculated with seacarb (`r round(const_choice[1, 6], digits = 3)`. In general, pH is higher with Sulpis than with Lueker.
- very good for CT (`r round(diff_ls_quant[4, 3], digits = 1)` umol/kg on average; always lower with Sulpis than with Lueker) compared with the uncertainty calculated with seacarb (`r round(const_choice[1, 8], digits = 1)`)
- negligible for OmegaAragonite

@@

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using various formulations of K1 and K2. All pH values were normalized to a temperature of 4°C. The absolute differences are large, up to `r round(max(abs(diff_table[, 2:5])), digits=3)` pH units. On average, both formulations overestimate spectrophotometric pH. The calculated pH closer to measured pH is using the formulation of Lueker et al. (2020) with a mean difference of `r round(max(diff_table[4, 2:5]), digits=3)` pH units. This difference is twice larger than the uncertainty for pH calculated from pCO2 and AT according to seacarb (`r round(errors_pco2_at$pH, digits=3)`).
`r knitr::kable(diff_table, digits = 3, caption = "Difference between spectrophotometric pH and pH calculated with pCO2 and AT using different formulations for K1 and K2.", format = "simple")`
`r include_graphics("figures/lueker-sulpis.png")`

>>>>

# Conversion of spec pH of in situ samples to in situ temp and 4 °C
pH_seafet <- d %>%
  dplyr::filter(!is.na(ph_s_sf) & !is.na(temp_insitu_ctd))
pH_seafet <- pH_seafet %>% 
  dplyr::mutate(
    ph_s_sf_temp_insi = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = pH_seafet$temp_insitu_ctd, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_sf_temp_4 = pHinsi(pH = pH_seafet$ph_s_sf, 
                         ALK = pH_seafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_seafet$temp_s_lab, Pinsi = pH_seafet$pressure_insitu_ctd/10, 
                         S = pH_seafet$sal_mix, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
  )
# Conversion of spec pH of FerryBox samples to FerryBox temp
pH_durafet <- d %>%
  tidyr::drop_na(ph_s_dur, sal_fb, temp_fb)
pH_durafet <- pH_durafet %>% 
  dplyr::mutate(
    ph_s_dur_temp_fb = pHinsi(pH = pH_durafet$ph_s_dur, 
                              ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = pH_durafet$temp_fb, 
                              Tlab = pH_durafet$temp_s_lab, Pinsi = 0, S = pH_durafet$sal_fb, 
                              Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74"),
   ph_s_dur_temp_4_l = pHinsi(pH = pH_durafet$ph_s_dur, # lueker
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_s20 = pHinsi(pH = pH_durafet$ph_s_dur, # Sulpis
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),       # Millero 02
   ph_s_dur_temp_4_p18 = pHinsi(pH = pH_durafet$ph_s_dur, # Papadimitriou
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74"),
   ph_s_dur_temp_4_m02 = pHinsi(pH = pH_durafet$ph_s_dur,  # Millero 02
                         ALK = pH_durafet$at_mean_interp*1e-6, Tinsi = 4, 
                         Tlab = pH_durafet$temp_s_lab, Pinsi = 0, 
                         S = pH_durafet$sal_fb, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")

   )
# Make mean for replicates + SD
pH_seafet_mean <- pH_seafet %>%
  dplyr::select(date, ph_s_sf_temp_insi, ph_s_sf_temp_4) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarise(ph_s_sf_temp_insi_mean = mean(ph_s_sf_temp_insi, na.rm = TRUE),
                   #ph_s_sf_temp_insi_sd = sd(ph_s_sf_temp_insi, na.rm = TRUE),
                   ph_s_sf_temp_4_mean = mean(ph_s_sf_temp_4, na.rm = TRUE)
                   )
pH_durafet_mean <- pH_durafet %>%
  dplyr::select(date, ph_s_dur_temp_fb, ph_s_dur_temp_4_l, ph_s_dur_temp_4_s20,
                ph_s_dur_temp_4_p18, ph_s_dur_temp_4_m02) %>% 
  dplyr::group_by(date) %>% 
  dplyr::summarize(ph_s_dur_temp_fb_mean = mean(ph_s_dur_temp_fb, na.rm = TRUE),
                   #ph_s_dur_temp_fb_sd = sd(ph_s_dur_temp_fb, na.rm = TRUE),
                   ph_s_dur_temp_4_l_mean = mean(ph_s_dur_temp_4_l, na.rm = TRUE),
                   ph_s_dur_temp_4_s20_mean = mean(ph_s_dur_temp_4_s20, na.rm = TRUE),
                   ph_s_dur_temp_4_p18_mean = mean(ph_s_dur_temp_4_p18, na.rm = TRUE),
                   ph_s_dur_temp_4_m02_mean = mean(ph_s_dur_temp_4_m02, na.rm = TRUE)
                   )

@@

# pH calculated (at and ct) at Ferrybox sal and temp
discrete_no_na <- discrete %>%
  dplyr::filter(!is.na(sal_fb)  & !is.na(temp_fb) & !is.na(at) & !is.na(ct))
discrete_is_na <- discrete %>%
  dplyr::filter(is.na(sal_fb)  | is.na(temp_fb) | is.na(at) | is.na(ct))
carb <- carb(15, discrete_no_na$at*1e-6, discrete_no_na$ct*1e-6, S = discrete_no_na$sal_fb, T = discrete_no_na$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
discrete_no_na$ph_calc_fb <- carb$pH # pH to compare with durafet pH (at Ferrybox temp and sal)
discrete_no_na$pco2_calc_fb <- carb$pCO2

# create new variable in "discrete_is_na" to bind "discrete_no_na"
discrete_is_na$ph_calc_fb <- NA
discrete_is_na$pco2_calc_fb <- NA
discrete <- bind_rows(discrete_no_na, discrete_is_na) %>% 
  dplyr::arrange(date)
```
```{r read and prepare nutrient data, echo=FALSE, warning=TRUE, message=FALSE}
# File "Ferrybox nutrients 2014-2018.xlsx" provided by 
# Willem van de Poll <w.h.van.de.poll@rug.nl> on 2020-01-16
# nutrients in umol/l
nuts <- readxl::read_excel(paste0(path, "fb_data/Ferrybox nutrients 2014-2018.xlsx"),
                           skip =1) %>% 
  drop_na(date) %>% 
  dplyr::mutate(datetime = as.POSIXct(paste0(date, " 12:00:00")),
                ) %>% # assuming sampling time 12:00
  dplyr::filter(datetime > as.POSIXct("2015-07-09"))
# The nearest times
time_index <- unique(d_all$datetime)
# nuts$closest <- time_index[
#   FNN::knnx.index(time_index, # The values to be queried (big df)
#                   as.vector(nuts$datetime), # The value to find the closest match to (small df)
#                   k = 1)] # The number of closest values to return
nuts <- nuts %>% 
  dplyr::mutate(closest = time_index[
  FNN::knnx.index(time_index, # The values to be queried (big df)
                  as.vector(nuts$datetime), # The value to find the closest match to (small df)
                  k = 1)]) %>%  # The number of closest values to return
  dplyr::select(-date, -datetime) %>% 
  dplyr::rename(NO3NO2 = `NO3+NO2`)
# cleanup and convert in umol/kg
d_all <- left_join(d_all, nuts, by = c("datetime" = "closest")) #%>% 
  #dplyr::select(-closest_datetime)
d_all <- d_all %>% 
  dplyr::mutate(density_fb = seacarb::rho(S = sal_fb, T = temp_fb, P = 0),
                po4 = PO4 / (density_fb/1000),
                nh4 = NH4 / (density_fb/1000),
                no3no2 = NO3NO2 / (density_fb/1000),
                no3 = NO3 / (density_fb/1000),
                si = Si / (density_fb/1000)
  )
```

@@

# pH calculated ferrybox
discrete2merge <- discrete %>%
  #drop_na() %>%
  dplyr::select(closest_datetime, ph_calc_fb) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

# durafet pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_dur_temp_fb_mean, ph_s_dur_temp_4_l_mean,
                ph_s_dur_temp_4_s20_mean, ph_s_dur_temp_4_p18_mean, ph_s_dur_temp_4_m02_mean) %>%
  dplyr::rename(ph_s_dur_temp_fb = ph_s_dur_temp_fb_mean,
                ph_s_dur_temp_4 = ph_s_dur_temp_4_l_mean, # lueker by default
                ph_s_dur_temp_4_l = ph_s_dur_temp_4_l_mean,
                ph_s_dur_temp_4_s20 = ph_s_dur_temp_4_s20_mean,
                ph_s_dur_temp_4_p18 = ph_s_dur_temp_4_p18_mean,
                ph_s_dur_temp_4_m02 = ph_s_dur_temp_4_m02_mean,) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")

@@

### **Calculated / Durafet pHs**
```{r pHs comparaison dur calc, echo=FALSE, message= FALSE, warning=FALSE}
# Normalize ph_calc from AT/CT to 4°c
pH_calc <- discrete %>%
  dplyr::select(datetime, sal_fb, temp_fb, ph_calc_fb, at)%>%
  dplyr::filter(!is.na(ph_calc_fb), !is.na(temp_fb), !is.na(sal_fb))
pH_calc$pH_calc_fb_temp_4 <- pHinsi(pH = pH_calc$ph_calc_fb, ALK=pH_calc$at*1e-6, Tinsi=4, Tlab=pH_calc$temp_fb, Pinsi=11/10, S=pH_calc$sal_fb, Pt=0, Sit=0, k1k2 = "l", kf = "dg", ks="d", pHscale = "T", b="u74")

@@

## Impact K1K2  {.tabset .tabset-pills}
### Lueker vs Sulpis
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
mean_discrete <- tibble(mean_ct = mean_ct, mean_at = mean_at, mean_pco2 = mean_pco2, mean_sal = mean_sal, mean_temp = mean_temp)

# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
Using the following using mean discrete values:
`r knitr::kable(mean_discrete, digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

@@

### K1K2 using pCO2 and AT
```{r impact of constants using FerryBox pCO2 and at, echo=FALSE, warning=FALSE, message=FALSE}

@@

  dplyr::select(datetime, ph_s_dur_temp_4_l, ph_s_dur_temp_4_s20,
                ph_s_dur_temp_4_p18, ph_s_dur_temp_4_m02, pco2_fb, at, sal_fb, temp_fb) %>% 
# This is an important adjustement for pco2_fb on 2019-06-07 09:00:00
# it is too high compared with ph_s_dur. Looking at the dygraph suggests that there has been a peak just before that time and that the values were coming back down. This is this in accordance with pco2_calc_fb calculated from 
dplyr::mutate(pco2_fb = ifelse(datetime == as.POSIXct("2019-06-07 09:00:00"), 203, pco2_fb))

@@

tmp2 <- tibble(datetime = tmp$datetime, ph_s_dur_temp_4_l = tmp$ph_s_dur_temp_4_l, 
               ph_s_dur_temp_4_s20 = tmp$ph_s_dur_temp_4_s20,
               ph_s_dur_temp_4_p18 = tmp$ph_s_dur_temp_4_p18,
               ph_s_dur_temp_4_m02 = tmp$ph_s_dur_temp_4_m02, 

@@

  tidyr::drop_na(ph_s_dur_temp_4_l)

@@

lueker <- as.numeric(summary(tmp2$ph_s_dur_temp_4_l - tmp2$pH_l))
sulpis <- as.numeric(summary(tmp2$ph_s_dur_temp_4_s20 - tmp2$pH_s))
papadimitriou <- as.numeric(summary(tmp2$ph_s_dur_temp_4_p18 - tmp2$pH_p))
millero2002 <- as.numeric(summary(tmp2$ph_s_dur_temp_4_m02 - tmp2$pH_m))

@@

p1 <- ggreg(xdata = tmp2$ph_s_dur_temp_4_l, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 

@@

            xlab = "Spectrophotometric pH normalized @ 4°C",

@@

p2 <- ggreg(xdata = tmp2$ph_s_dur_temp_4_l, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 

@@

  dplyr::mutate(delta_lueker = tmp2$ph_s_dur_temp_4_l - tmp2$pH_l, 
                delta_sulpis = tmp2$ph_s_dur_temp_4_s20 - tmp2$pH_s, 
                delta_papadimitriou = tmp2$ph_s_dur_temp_4_p18 - tmp2$pH_p, 
                delta_millero2002 = tmp2$ph_s_dur_temp_4_m02 - tmp2$pH_m) %>% 

@@

ggsave(filename = "figures/k1k2_all_pco2_at.png", plot = p, width = 25, height = 20, units = "cm")

# save elements for k1k2 report
k1k2_pco2_at <- list(mean_discrete = mean_discrete, const_choice = const_choice, tmp2 = tmp2,
             errors_pco2_at = errors_pco2_at, diff_table = diff_table, 
             diff_ls_quant = diff_ls_quant)
save(k1k2_pco2_at, file="reports/k1k2_pco2_at.RData")
```

The discrete values of pCO2, AT, salinity and temperature in the FerryBox were used to calculate pH using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants are:

- large for pH (`r round(diff_ls_quant[4, 2], digits = 3)` pH units on average; ; median is `r round(diff_ls_quant[3, 2], digits = 3)`) compared with the uncertainty calculated with seacarb (`r round(const_choice[1, 6], digits = 3)`). In general, pH is higher with Sulpis than with Lueker.
- small for CT (`r round(diff_ls_quant[4, 3], digits = 1)` umol/kg on average; always lower with Sulpis than with Lueker) compared with the uncertainty calculated with seacarb (`r round(const_choice[1, 8], digits = 1)`)
- negligible for OmegaAragonite

@@

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using various formulations of K1 and K2. All pH values were normalized to a temperature of 4°C. The absolute differences are generally large, up to `r round(max(abs(diff_table[, 2:5])), digits=3)` pH units. In general, all formulations overestimate spectrophotometric pH. pH calculated using the formulation of Papadimitriou et al. (2018) is closer to measured pH, with a mean difference of `r round(diff_table[4, 4], digits=3)` pH units. This difference is `r abs(round(diff_table[4, 4]/errors_pco2_at$pH, digits=1))` times larger than the uncertainty for pH calculated from pCO2 and AT according to seacarb (`r round(errors_pco2_at$pH, digits=3)`).
`r knitr::kable(diff_table, digits = 3, caption = "Difference between spectrophotometric pH and pH calculated with pCO2 and AT using different formulations for K1 and K2.", format = "simple")`
`r include_graphics("figures/k1k2_all_pco2_at.png")`

### K1K2 using CT and AT
```{r impact of constants using FerryBox ct and at, echo=FALSE, warning=FALSE, message=FALSE}
  tidyr::drop_na(ct, at, sal_fb, temp_fb) %>% 
  dplyr::select(datetime, ph_s_dur_temp_4_l, ph_s_dur_temp_4_s20,
